---
title: "Kernel-Density-Estimation and Bandwidth Estimation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## General
Load the package with: 
```{r setup}
library(KDE)
```

## Sampling

## Kernel-Density-Estimation
```
kernel_density_estimator(kernel, samples, bandwidth=1, subdivisions=100L)
```
## Bandwidth Estimation

Explenation/Purpose of 

* Kernel 
* Samples
* Bandwidths
* kappa/lambda
* subdivisions

### Cross-Validation

```
cross_validation(kernel, samples, bandwidths = logarithmic_bandwidth_set(1/length(samples), 1, 10), subdivisions = 100L)
```

### Goldenshluger-Lepski Method

```
goldenschluger_lepski <- function(kernel, samples, bandwidths = logarithmic_bandwidth_set(1/length(samples), 1, 10), kappa = 1.2, subdivisions = 100L)
```

### Penalized-Comparison-to-Overfitting Method

To get the optimal bandwidth from the given bandwidth set
```
pco_method <- function(kernel, samples, bandwidths = NULL, lambda = 1, subdivisions = 100L)
```

To get a list of the bandwidths and their associated risk values

```
pco_crit <- function(kernel, samples, bandwidths = logarithmic_bandwidth_set(1/length(samples), 1, 10), lambda = 1, subdivisions = 100L)    ```                 
